Supporting Non-Functional Requirements in Services Software Development Process: An MDD Approach

### 2013-10-21 SOFSEM2014/FinalVersion/pisodm.tex
New figure and more modifications maintaining 12 pages
Changes in the sodm.tex and pisodm.tex, The included text are marked with comments.

\begin{figure}[h]
\centering
\includegraphics[width=1.0\textwidth]{figs/piSODM}
\caption{$\pi$SOD-M.}
\label{fig:piSOD-M}
\end{figure}

@@

%
%Similarly to SOD-M, our approach targets the construction of service-oriented applications that implement business processes.
Thus, $\pi$SOD-M proposes a development process based on the definition of models
(instances of the meta-models) and transformations between models.
There are two kinds of transformations:
Model-to-model transformations are used during the software process to refine the specification.
Model-to-text transformations are the last step of the process and generate code.
In the next section we develop an example, to serve as a proof-of-concept.
The example will show the actual notation used for models.

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{figs/piSOD-M.pdf}
\caption{$\pi$SOD-M.}
\label{fig:piSOD-M}
\end{figure}

@@

%
%Similarly to SOD-M, our approach targets the construction of service-oriented applications that implement business processes.
Thus, $\pi$SOD-M proposes a development process based on the definition of models
(instances of the meta-models) and transformations between models.
There are two kinds of transformations:
Model-to-model transformations are used during the software process to refine the specification.
Model-to-text transformations are the last step of the process and generate code.

%%%BEGIN{Text Included by Placido Neto for Camera-Ready}%%%%%%
$\pi$SOD-M environment is built on the top of Eclipse framework, which is a
framework to build Integrated Development Environments (IDEs). We also use the
Eclipse Modelling Framework (EMF), a meta-modelling framework that was devised to be
extended and provides with the utilities needed to define, edit and handle
(meta)-models. To automate the transformation models we use the ATL Language\footnote{ATL uses EMF to handle models (that is to serialize, to navigate and to modify
them). Once the ATL transformation is executed, the ATL engine matches the
source pattern and the source model and instantiates each pattern into the
target model.} \cite{atl_manual}, a model transformation language framed in
Eclipse. Another language used for model transformation is Acceleo \cite{acceleo}, a
model to text engine. $\pi$SOD-M uses this environment~\footnote{The environmental requirements are the Eclipse Modelling (Galilleo or higher)
and all proposed plugins, \textit{$\pi$-UseCase, $\pi$-ServiceProcess,
$\pi$-SerciveComposition} and \textit{$\pi$-PEWS} editors; and also the
transformation plugins. Because the environment is proposed in the
context of Eclipse, the environmental requirements are also the same needed to
utilize the Eclipse tool.} to generate service
composition specification code in \textit{$\pi$-PEWS} and semi-automatize the
 implementation of reliable service-based applications.
%%%END{Text Included by Placido Neto for Camera-Ready}%%%%%%

In the next section we develop an example, to serve as a proof-of-concept.
The example will show the actual notation used for models.

### 2013-10-21 SOFSEM2014/FinalVersion/sodm.tex
New figure and more modifications maintaining 12 pages
Changes in the sodm.tex and pisodm.tex, The included text are marked with comments.

The SOD-M approach includes transformations between models:
\textit{CIM-to-PIM, PIM-to-PIM} and \textit{PIM-to-PSM} transformations. Given
an abstract model at the CIM level, it is possible to apply transformations for
generating a model of the PSM level. In this context, it is necessary to
follow the process activities described by the methodology.

These three SOD-M levels have no support for describing non-functional requirements.The following section introduces $\pi$-SODM the extension proposed for considering these requirements.

>>>>

 %%%BEGIN{Text Included by Placido Neto for Camera-Ready}%%%%%%
The \textit{value model} is a business model that describes a business
case as a set of values and value activities shared by business actors. The
\textit{BPMN model} (business process model) is used to describe the business
process related to the environment which the system will run. These two models
represent the independent aspects of computing. The \textit{use case model} is used to represent
the business services to be implemented by the system, while the \textit{extended use case model} is a
behavioral model, to represent the system features as a way to implement
the business services. The \textit{service process model} describes the set of
activities that must be performed on the system to implement a business service.
Finally, the \textit{service composition model} represents the full flow of
business system. This model is an extension of the service process model,
however, in more detail. These four models represent the platform independent
aspects.
%%%END{Text Included by Placido Neto for Camera-Ready}%%%%%%

The SOD-M approach includes transformations between models:
\textit{CIM-to-PIM, PIM-to-PIM} and \textit{PIM-to-PSM} transformations. Given
an abstract model at the CIM level, it is possible to apply transformations for
generating a model of the PSM level. In this context, it is necessary to
follow the process activities described by the methodology. 
%
These three SOD-M levels have no support for describing non-functional requirements.The following section introduces $\pi$-SODM the extension proposed for considering these requirements.
